Quantum Magic Games

### 2023-07-04 paper/main.tex
paper => intro

\usepackage[parfill]{parskip} % noindent + space between paragraph (I really prefer...)
\usepackage[table]{xcolor}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{nicefrac}
\usepackage{hhline}
\usepackage{bigdelim}
\usepackage{booktabs}

@@

\maketitle
% \begin{abstract}
% \end{abstract}

@@

Telepathy, the ability of transmitting information from one
person's mind to another's, would certainly come in handy in
many situations, right? Unfortunately (or not), (to the best
of our knowledge) telepathy is not a thing. At least, not
according to classical physics. Certain aspects of the
quantum realm, however, provide a way of communication that
for a layman looks as magical as ``true'' telepathy. This
phenomenon is called quantum
\emph{pseudo-telepathy}~\cite{brassard:2005}.

Quantum pseudo-telepathy is observed in many contexts,
usually described in the format of a game: the ``impossible
colouring games''~\cite{brassard:2005, Kochen1975}; the
parity games, in which \(n \geq 3\) players are given
bit-strings and, without communicating to each other, they
output one of their bits, winning if their outputs combined
obey certain parity conditions~\cite{brassard:2005,
  Mermin1990}; the Deutsch-Jozsa games, where Alice and Bob
are given bit strings \(x\) and \(y\), and must output bit
strings \(a\) and \(b\) such that \(a = b\) if and only if
\(x = y\)~\cite{brassard:2005, Brassard_1999}; and, the
Magic Square game~\cite{brassard:2005, Mermin1990}. None of
these games admit a classical winning strategy (i.e.\ is not
possible to always win), yet they can be won systematically,
without any communication, provided that the players share
prior entanglement~\cite{brassard:2005}.

In this project, we shall explore the Mermin-Peres Magic Square game
(hereinafter referred simply as Magic Square). The origins of this
game date back to the ninities. It was first described ---~albeit not
in the format of a game~--- in the works of Mermin~\cite{mermin:1990}
and of Peres~\cite{peres:1990}. Their results provided (as per the
title of Mermin's review letter) a simplied proof for the
Kochen-Specker theorem~\cite{Kocher1975}. Later, in 2002, Aravind
demonstrated how to transform Mermin-Peres's proof of then
Kochen-Specker theorem into a proof of Bell's
theorem~\cite{aravind:xxx}. Bell's theorem shows the incompatibility
of hidden variables (i.e. determinism) and locality, whereas
Kochen-Specker's theorem demonstrates the conflicting nature of
determinism and non-contextuality.  These sort of results may seem
challenging, but their descriptions can be greatly simplified by
modeling them as quantum games.  In particular, it is extremely easy
to show that there cannot be a classical solution to the the Magic
Square game, and to convince an observer that something ``magical''
(classically impossible) is happening in a successful implementation
of a quantum winnin strategy~\cite{brassard:2005}.

\todo{FIX: they talk about contextuality only (both theorems)}

\paragraph{Outline of this paper.}

@@

>>>>

\usepackage[hidelinks]{hyperref}
\usepackage[parfill]{parskip} % noindent + space between paragraph (I really prefer...)
\usepackage[table]{xcolor}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{nicefrac}
\usepackage{hhline}
\usepackage{bigdelim}
\usepackage{booktabs}

@@

\maketitle
\begin{abstract}
  \textbf{TODO}
\end{abstract}

@@

Telepathy, the ability of transmitting information from one
person's mind to another's, would certainly come in handy in
many situations, right? Unfortunately (or not), (to the best
of our knowledge) telepathy is not a thing. At least, not
according to classical physics. Certain aspects of the
quantum realm, however, provide a way of communication that,
for a layman, may look as ``magical'', or as true telepathy. This
phenomenon is called quantum
\emph{pseudo-telepathy}~\cite{brassard:2005}.

Quantum pseudo-telepathy is observed in many contexts, usually
described in the format of a game. Examples of such games are: the
``impossible colouring games''~\cite{cleve:2004}, based on the
Kochen-Specker theorem~\cite{kochen:1967};
%% Avoid confusion with our definition of parity telepathy games
% the
% ``parity games'', in which \(n \geq 3\) players are given
% bit-strings and, without communicating to each other, they
% output one of their bits, winning if their outputs combined
% obey certain parity conditions~\cite{brassard:2005,
%   Mermin1990};
the ``Deutsch-Jozsa games''~\cite{brassard:1999}, based on the Deutsch-Jozsa problem~\cite{deutsch:1992};
% where Alice and Bob
% are given bit strings \(x\) and \(y\), and must output bit
% strings \(a\) and \(b\) such that \(a = b\) if and only if
% \(x = y\)~\cite{brassard:2005};
the ``magic square'' game~\cite{aravind:2002}, based on the Mermin-Peres proof of the
Kochen-Specker theorem~\cite{mermin:1990, peres:1990, mermin:1993, peres:1997}; and many
others~\cite{cleve:2004, brassard:2005}. None of these games admit a
classical winning strategy (i.e.\ is not possible to always win), yet
they can be won systematically, without any communication, provided
that the players are allowed to share prior entanglement~\cite{brassard:2005}.

In this project, we will explore the Mermin-Peres magic square game
(hereinafter referred simply as magic square). The origins of this
game date back to the ninities. It was first described in the works of
Mermin~\cite{mermin:1990, mermin:1993} and of
Peres~\cite{peres:1990, peres:1997}. Their results provided (as per the title of
Mermin's review letter) a simplied proof for the Kochen-Specker
theorem~\cite{kochen:1967}. Later, in 2002, Aravind demonstrated how
to transform Mermin-Peres's proof of the Kochen-Specker theorem into a
proof of Bell's theorem~\cite{aravind:2002}. Bell's
theorem~\cite{bell:1964} shows the incompatibility of hidden variables
(i.e. determinism) and locality, and Kochen-Specker's theorem
demonstrates the conflicting nature of determinism and
noncontextuality (locality can be regarded as a special case of
contextuality). These sort of results may seem
challenging, but their descriptions can be greatly simplified by
modeling them as quantum games.  In particular, it is extremely easy
to show that there cannot be a classical solution to the the Magic
Square game, and to convince an observer that something ``magical''
(classically impossible) is happening in a successful implementation
of a quantum winnin strategy~\cite{brassard:2005}.

\paragraph{Outline of this paper.}
In Section~\ref{sec:magic-square}, we detail the intrinsics of the
magic square game, classical approaches and why it is impossible to
systematically win the game classically, how and why the quantum
strategy works, and how to implement it. We also briefly discuss how
the game translates to the proof of the incompatibility of a
noncontextual hidden-variable theory. Then, in
Section~\ref{sec:extending}, we describe a generalization of games
that follow the same style of the magic square game, given
by Arkhipov~\cite{arkhipov:2012}. Finally, section~\ref{sec:conc} concludes the
presentation.

\paragraph{Code.} The implementation of the classical and quantum
solutions to the magic square game can be found at
\url{https://github.com/luigidcsoares/quantum-magic-square}.

@@

\section{Conclusion}\label{sec:conc}
